### 2023-04-12 chapters/paper_benchmarking_workshop.tex
Add review paper (WIP)
Change citation keys (WIP)
Update reference keys for review paper

\section{Introduction}
Since their introduction, temporal latent variable models (LVMs) for speech \cite{chung_recurrent_2015, fraccaro_sequential_2016} based on the variational autoencoder (VAE,  \textcite{kingma_auto-encoding_2014, rezende_stochastic_2014}) have seen little development compared to their image domain counterparts. While LVMs now achieve superior likelihoods on images compared to deterministic models \cite{child_very_2021,sinha_consistency_2021,kingma_variational_2021}, they are still inferior to deterministic models such as WaveNet on speech \cite{oord_wavenet_2016}.
Development in the image domain has been driven by established likelihood benchmarks, but in the speech domain, likelihoods are often not reported \cite{oord_wavenet_2016, hsu_unsupervised_2017, oord_neural_2018} or are incomparable due to subtle differences in the chosen data distributions \cite{chung_recurrent_2015, fraccaro_sequential_2016, hsu_unsupervised_2017, aksan_stcn_2019}. 
This makes it hard to develop explicit likelihood models for speech.

@@

        \edge[]{d_t}{z_t};
        \edge[bend right]{d_t}{x_t};
        \edge[]{z_tm1}{d_t};
        \edge[]{d_tm1}{d_t};
        \edge[]{z_t}{x_t};
    \end{tikzpicture}
    }
    \end{multicols}
    \vspace{-4mm}
\caption{
Generative models of an LSTM, VRNN, SRNN, STCN and CW-VAE for a single time step. The STCN and CW-VAE are illustrated with a single latent variable. 
Red arrows indicate purely deterministic paths from the output $\vx_t$ to previous input $\vx_{<t}$ without passing a stochastic node. 
We provide additional graphical illustrations including inference models in appendix~\ref{app: additional graphical models}.
}
\label{fig: autoregressive vrnn srnn cwvae 1L graphs}
\vspace{-3mm}
\end{figure*}

@@

\section{Models}
\paragraph{Sequential deep latent variable models}
The models we consider are all sequential deep latent variable models trained with variational inference and the reparameterization trick \cite{kingma_auto-encoding_2014}.
The input is a variable-length sequence $\vx = \vx_{1:T} = \left(\vx_1, \vx_2, \dots, \vx_{T}\right)$ with $\vx_t\in\mathbb{R}^{D_x}$. 
The models first encode $\vx_{1:T}$ to a temporal posterior distribution $q(\vz|\vx)$ and sample a latent representation $\vz_{1:T}$ where $\vz_t\in\mathbb{R}^{D_z}$. This is then used to reconstruct the input while the posterior distribution is regularized to be close to a prior distribution $p(\vz_t|\cdot)$ where the dot indicates that it may depend on latent and observed variables at previous time steps, $\vz_{<t}$ and $\vx_{<t}$ where $\vz_{<t} := \left( \vz_1, \vz_2, \dots, \vz_{t-1} \right)$. 
The models are trained to maximize a lower bound $\mathcal{L}(\thetab,\phib;\vx)$ on the marginal likelihood $\log p_\thetab(\vx)$,
\begin{equation}
    \log p_\thetab(\vx) = \log \int p_\thetab(\vx,\vz) \, \mathrm{d}\vz  \geq \int q_\phib(\vz|\vx) \log \frac{p_\thetab(\vx|\vz)p(\vz)}{q_\phib(\vz|\vx)}  \, \mathrm{d}\vz \triangleq \mathcal{L}(\thetab,\phib;\vx)  \enspace , \label{eq: elbo}
\end{equation}
where $\thetab$ are parameters of the generative model and $q_\phib(\vz|\vx)$ is the variational approximation to the true posterior.
Graphical illustrations of the models can be seen in \cref{fig: autoregressive vrnn srnn cwvae 1L graphs} and appendix~\cref{app: additional graphical models}.

>>>>

\section{Introduction}
Since their introduction, temporal latent variable models (LVMs) for speech \cite{chung_recurrent_2015, fraccaro_sequential_2016} based on the variational autoencoder (VAE,  \textcite{kingma_autoencoding_2014, rezende_stochastic_2014}) have seen little development compared to their image domain counterparts. While LVMs now achieve superior likelihoods on images compared to deterministic models \cite{child_very_2021,sinha_consistency_2021,kingma_variational_2021}, they are still inferior to deterministic models such as WaveNet on speech \cite{oord_wavenet_2016}.
Development in the image domain has been driven by established likelihood benchmarks, but in the speech domain, likelihoods are often not reported \cite{oord_wavenet_2016, hsu_unsupervised_2017, oord_neural_2018} or are incomparable due to subtle differences in the chosen data distributions \cite{chung_recurrent_2015, fraccaro_sequential_2016, hsu_unsupervised_2017, aksan_stcn_2019}. 
This makes it hard to develop explicit likelihood models for speech.

@@

        \edge[]{d_t}{z_t};
        \edge[bend right]{d_t}{x_t};
        \edge[]{z_tm1}{d_t};
        \edge[]{d_tm1}{d_t};
        \edge[]{z_t}{x_t};
    \end{tikzpicture}
    }
    \end{multicols}
\caption{
Generative models of an LSTM, VRNN, SRNN, STCN and CW-VAE for a single time step. The STCN and CW-VAE are illustrated with a single latent variable. 
Red arrows indicate purely deterministic paths from the output $\vx_t$ to previous input $\vx_{<t}$ without passing a stochastic node. 
We provide additional graphical illustrations including inference models in \cref{app: additional graphical models}.
}
\label{fig: autoregressive vrnn srnn cwvae 1L graphs}
\end{figure*}

@@

\section{Models}
\paragraph{Sequential deep latent variable models}
The models we consider are all sequential deep latent variable models trained with variational inference and the reparameterization trick \cite{kingma_autoencoding_2014}.
The input is a variable-length sequence $\vx = \vx_{1:T} = \left(\vx_1, \vx_2, \dots, \vx_{T}\right)$ with $\vx_t\in\mathbb{R}^{D_x}$. 
The models first encode $\vx_{1:T}$ to a temporal posterior distribution $q(\vz|\vx)$ and sample a latent representation $\vz_{1:T}$ where $\vz_t\in\mathbb{R}^{D_z}$. This is then used to reconstruct the input while the posterior distribution is regularized to be close to a prior distribution $p(\vz_t|\cdot)$ where the dot indicates that it may depend on latent and observed variables at previous time steps, $\vz_{<t}$ and $\vx_{<t}$ where $\vz_{<t} := \left( \vz_1, \vz_2, \dots, \vz_{t-1} \right)$. 
The models are trained to maximize a lower bound $\mathcal{L}(\thetab,\phib;\vx)$ on the marginal likelihood $\log p_\thetab(\vx)$,
\begin{equation}
    \log p_\thetab(\vx) = \log \int p_\thetab(\vx,\vz) \, \mathrm{d}\vz  \geq \int q_\phib(\vz|\vx) \log \frac{p_\thetab(\vx|\vz)p(\vz)}{q_\phib(\vz|\vx)}  \, \mathrm{d}\vz \triangleq \mathcal{L}(\thetab,\phib;\vx)  \enspace , \label{eq: elbo}
\end{equation}
where $\thetab$ are parameters of the generative model and $q_\phib(\vz|\vx)$ is the variational approximation to the true posterior.
Graphical illustrations of the models can be seen in \cref{fig: autoregressive vrnn srnn cwvae 1L graphs} and appendix~\cref{app: additional graphical models}.

### 2023-04-12 chapters/paper_hierarchical.tex
Change citation keys (WIP)
Update reference keys for review paper

\section{Why does OOD detection fail?}\label{sec:why-does-ood-fail}
\begin{figure}[t]
    \centering
    \includegraphics[width=1\columnwidth]{paper_hierarchical/feature-correlation-heatmap2.pdf}
    \vspace{-5mm}
    \caption[Absolute correlations between data representations in all layers of the inference network of a hierarchical VAE trained on FashionMNIST and of another trained on MNIST.]{
    Absolute correlations between data representations in all layers of the inference network of a hierarchical VAE trained on FashionMNIST and of another trained on MNIST.
    We compute the correlation between the representations of the two different models given the same data, FashionMNIST (top) and MNIST (bottom).
    %We reduce the representations to a single number by summing over all dimensions and compute the correlation between these numbers for 256 examples.
    %To reduce de-correlation due to the stochastic latent variables, we encode using the mode of the approximate posterior distributions.
    }
    \vspace{-3mm}
    \label{fig:correlations-heatmap}
\end{figure}

@@

\begin{figure}[t]
    \centering
    %\vspace{0.17cm}
    \includegraphics[width=1\columnwidth]{paper_hierarchical/biva_reconstructions.pdf}
    % https://docs.google.com/drawings/d/1UXdA3c18Oek9_env1vEB9mqOAHZ-tu0ijWmvqmapWCk/edit?usp=sharing
    \vspace{-5.2mm}
    \caption[Reconstructions of in-distribution data (CelebA) of the BIVA model using higher latent variables]{Reconstructions of in-distribution data (CelebA) of the BIVA model using higher latent variables  \cite{maaloe_biva_2019}.
    The higher the latent variable, the more the reconstructions fall into the mode of the learned distribution.
    It is more common to wear regular glasses than sunglasses but most common not to wear glasses at all.
    A man with long hair collapses into the mode of the more common long-haired woman.}
    \vspace{-3mm}
    \label{fig:biva-reconstructions-celeba}
\end{figure}
Finally, we look at reconstructions of out-of-distribution data.
\cref{fig:reconstructions-fashionmnist} illustrates that MNIST data is surprisingly well reconstructed by a hierarchical VAE trained on FashionMNIST.
Similar results have been found elsewhere \cite{xiao_likelihood_2020}.
We repeat the previous experiment and replace inference distributions by their corresponding conditional prior, and now observe that reconstructions from higher latent layers become increasingly similar to the data on which the model was trained.
The reliance on conditional priors seems to prevent accurate reconstruction of out-of-distribution data.
Some details are lost on in-distribution data too, but the distinction between that and out-of-distribution data becomes more clear.

@@

\vspace{-0.35cm}
\section{Background and related work}

@@

Hence, it is clear that compared to the likelihood bound $\mathcal{L}^{>k}$, this likelihood-ratio measures divergence exclusively in the latent space whereas $\mathcal{L}^{>k}$ includes the $\log p_\thetab(\xb)$ term similar to the ELBO.
Therefore, the $LLR^{>k}$ score should be an improved method for semantic OOD detection compared to $\mathcal{L}^{>k}$.
Now, it can be noted that if we replace the regular ELBO, $\mathcal{L}$, in \eqref{eq:likelihoods-as-exact} with the strictly tighter importance weighted bound \cite{burda_importance_2016},
\begin{equation}
    \mathcal{L}_{S} = \mathbb{E}_{q(\zb|\xb)}\left[ \log \frac{1}{N} \sum_{s=1}^{S} \frac{p(\xb, \zb^{(s)})}{q(\zb^{(s)}|\xb)} \right] \ , \label{eq:iw-bound}
\end{equation}
then, in the limit $S\rightarrow\infty$, we have $\mathcal{L}_{S} \rightarrow \log p_\thetab(\xb)$ and the likelihood ratio reduces to
\begin{equation}
    LLR^{>k}_{S}(\xb) \rightarrow D_{\mathrm{KL}}( p(\zb_{\leq k}|\zb_{>k}) q(\zb_{>k}|\xb) || p(\zb|\xb)) \label{eq:llr-as-kls-iwae-reduced}
\end{equation}
which, in practice, is well-approximated for a finite $S$. We expect this importance weighted likelihood ratio to monotonically improve upon the one in \cref{eq:llr-as-kls} as $S$ increases and the KL-divergence in the regular ELBO that contains terms for which $\zb_i$ is high-dimensional goes to zero.

Since the scores in \cref{eq:llr-as-kls, eq:llr-as-kls-iwae-reduced} are estimated by sampling their estimators are stochastic objects with nonzero variance.
We note that $\text{Var}(\widehat{LLR}^{>k}) = \text{Var}(\hat{\mathcal{L}}) + \text{Var}(\hat{\mathcal{L}}^{>k}) - 2\, \text{Cov}(\hat{\mathcal{L}}, \hat{\mathcal{L}}^{>k})$.
Since $\log p_\thetab(\xb)$ and part of the KL divergence are identical in the expressions of $\mathcal{L}$ and $\mathcal{L}^{>k}$ we expect $\text{Cov}(\hat{\mathcal{L}}, \hat{\mathcal{L}}^{>k})$ to be positive which reduces the total variance. 
Empirical results indeed show that $\text{Var}(\widehat{LLR}^{>k})$ is larger than $\text{Var}(\hat{\mathcal{L}})$ but smaller than $\text{Var}(\hat{\mathcal{L}}^{>k})$.
%Whether this decreases the variance to below the variance of any of the two bounds, we do not know, and we believe it to be difficult to verify mathematically.
Nevertheless, the variance of the estimators is guaranteed to go to zero as the number of samples is increased.

@@

% Our model
%\begin{wrapfigure}{R}{0.5\columnwidth}
% \begin{SCfigure}[50][t!]
\begin{figure}[t]
    % \vspace{-1.3cm}
    %\begin{figure}
    \centering
    \tikz{
        % inference
        % nodes
        \node[obs] (x_inf) {$\xb$};%
        \node[latent,above=.75cm of x_inf](z1_inf){$\zb_1$}; %
        \node[latent,above=.75cm of z1_inf](z2_inf){$\zb_2$}; %
        % \node[latent,above=.75cm of z2_inf](z3_inf){$\zb_3$}; %
        \node[above=of z2_inf, yshift=-1.cm] (phi) {$q_\phib(\zb|\xb)$};

@@

        % edges
        % \edge[]{z3_gen}{z2_gen};
        \edge[]{z2_gen}{z1_gen};
        \edge[]{z1_gen}{x_gen};
        \edge[dashed, bend left]{z2_gen}{x_gen};
        % \edge[dashed, bend left]{z3_gen}{x_gen};
    }
    \caption[Inference and generative models for a bottom-up hierarchical VAS]{The inference and generative models, $q_\phib$ and $p_\thetab$, for an $L=2$ layered bottom-up hierarchical VAE as the one used in our experiments.
    Dashed lines indicate deterministic skip connections which are employed in both networks. Skip connections are found to be useful for optimizing latent variable models \parencite{dieng_avoiding_2019, maaloe_biva_2019}.}
    \label{fig:hvae-graphical-model}
    % \end{figure}
    %\vspace{-0.8cm}
%\end{wrapfigure}
% \end{SCfigure}
\end{figure}

@@

\begin{table}[t!]
    \centering
    \resizebox{0.8\columnwidth}{!}{%
    \begin{tabular}{lrrrrr}
        \toprule
         Method & Dataset & \multicolumn{4}{c}{Avg. bits/dim}\\
        %   & & $\log p(x)$ & $\mathcal{L}^{>k_1}$ & $\mathcal{L}^{>k_2}$ & $\mathcal{L}^{>k_3}$\\
          & & $\log p(x)$ & $\mathcal{L}^{>1}$ & $\mathcal{L}^{>2}$ & $\mathcal{L}^{>3}$\\
         \midrule
         \multicolumn{6}{c}{\textbf{Trained on FashionMNIST}} \\
         \midrule
         \multirow{2}{*}{Glow}
            & FashionMNIST & 2.96 & - & - & \\
            & MNIST & 1.83 & - & - & \\
         \multirow{2}{*}{HVAE (Ours)}
            & FashionMNIST & 0.420 & 0.476 & 0.579 & - \\
            & MNIST & 0.317 & 0.601 & 0.881 & - \\
         \midrule
         \multicolumn{6}{c}{\textbf{Trained on CIFAR10}} \\
         \midrule
         \multirow{2}{*}{Glow}
          & CIFAR10 & 3.46 & - & - & \\
          & SVHN & 2.39 & - & - & \\
         \multirow{2}{*}{HVAE (Ours)}
            & CIFAR10 & 3.74 & 17.8 & 54.3 & 75.7 \\  % log p(x) = 8010.03
            & SVHN & 2.62 & 10.2 & 64.0 & 93.9 \\
         \multirow{2}{*}{BIVA (Ours)}
          & CIFAR10 & 3.46 & 8.74 & 19.7 & 37.3 \\
          & SVHN & 2.35 & 6.62 & 25.1 & 59.0 \\
         \bottomrule
    \end{tabular}
    }
    \caption[Average bits per dimension of different datasets for generative models trained on FashionMNIST and CIFAR10]{%
    Average bits per dimension of different datasets for models trained on FashionMNIST and CIFAR10.
    For the hierarchical models we include the $\mathcal{L}^{>k}$ bounds.
    The likelihoods of training and test splits of the in-distribution data are all cases close.
    Since we train on dynamically binarized FashionMNIST, our bits/dim are smaller than for Glow.
    As $k$ is increased for the $L^{>k}$ bound, the bound gets looser but the model eventually assigns higher likelihood to the in distribution data than to the OOD data.
    Glow refers to \textcite{kingma_glow_2018, nalisnick_deep_2019}.
    BIVA refers to our implementation of \textcite{maaloe_biva_2019}.}
    \label{tab:bits-per-dim-ood}
    \vspace{-3mm}
\end{table}

\footnotetext{

@@

    Performance is worse when the overlap is high, cf. \textcite[Table 1]{serra_input_2020}, as seen with complex images. 

>>>>

\section{Why does OOD detection fail?}\label{sec:why-does-ood-fail}
\begin{figure}[t]
    \centering
    \includegraphics[width=1\columnwidth]{paper_hierarchical/feature-correlation-heatmap2.pdf}
    \vspace{0mm}
    \caption[Absolute correlations between data representations in all layers of the inference network of a hierarchical VAE trained on FashionMNIST and of another trained on MNIST.]{
    Absolute correlations between data representations in all layers of the inference network of a hierarchical VAE trained on FashionMNIST and of another trained on MNIST.
    We compute the correlation between the representations of the two different models given the same data, FashionMNIST (top) and MNIST (bottom).
    %We reduce the representations to a single number by summing over all dimensions and compute the correlation between these numbers for 256 examples.
    %To reduce de-correlation due to the stochastic latent variables, we encode using the mode of the approximate posterior distributions.
    }
    \vspace{0mm}
    \label{fig:correlations-heatmap}
\end{figure}

@@

\begin{figure}[t]
    \centering
    %\vspace{0.17cm}
    \includegraphics[width=1\columnwidth]{paper_hierarchical/biva_reconstructions.pdf}
    % https://docs.google.com/drawings/d/1UXdA3c18Oek9_env1vEB9mqOAHZ-tu0ijWmvqmapWCk/edit?usp=sharing
    \vspace{0mm}
    \caption[Reconstructions of in-distribution data (CelebA) of the BIVA model using higher latent variables]{Reconstructions of in-distribution data (CelebA) of the BIVA model using higher latent variables  \cite{maaloe_biva_2019}.
    The higher the latent variable, the more the reconstructions fall into the mode of the learned distribution.
    It is more common to wear regular glasses than sunglasses but most common not to wear glasses at all.
    A man with long hair collapses into the mode of the more common long-haired woman.}
    \vspace{0mm}
    \label{fig:biva-reconstructions-celeba}
\end{figure}
Finally, we look at reconstructions of out-of-distribution data.
\cref{fig:reconstructions-fashionmnist} illustrates that MNIST data is surprisingly well reconstructed by a hierarchical VAE trained on FashionMNIST.
Similar results have been found elsewhere \cite{xiao_likelihood_2020}.
We repeat the previous experiment and replace inference distributions by their corresponding conditional prior, and now observe that reconstructions from higher latent layers become increasingly similar to the data on which the model was trained.
The reliance on conditional priors seems to prevent accurate reconstruction of out-of-distribution data.
Some details are lost on in-distribution data too, but the distinction between that and out-of-distribution data becomes more clear.

@@

\vspace{0cm}
\section{Background and related work}

@@

Hence, it is clear that compared to the likelihood bound $\mathcal{L}^{>k}$, this likelihood-ratio measures divergence exclusively in the latent space whereas $\mathcal{L}^{>k}$ includes the $\log p_\thetab(\xb)$ term similar to the ELBO.
Therefore, the $LLR^{>k}$ score should be an improved method for semantic OOD detection compared to $\mathcal{L}^{>k}$.
Now, it can be noted that if we replace the regular ELBO, $\mathcal{L}$, in \eqref{eq:likelihoods-as-exact} with the strictly tighter importance weighted bound \cite{burda_importance_2016},
\begin{equation}
    \mathcal{L}_{S} = \mathbb{E}_{q(\zb|\xb)}\left[ \log \frac{1}{N} \sum_{s=1}^{S} \frac{p(\xb, \zb^{(s)})}{q(\zb^{(s)}|\xb)} \right] \ , \label{eq:iw-bound}
\end{equation}
then, in the limit $S\rightarrow\infty$, we have $\mathcal{L}_{S} \rightarrow \log p_\thetab(\xb)$ and the likelihood ratio reduces to
\begin{equation}
    LLR^{>k}_{S}(\xb) \rightarrow D_{\mathrm{KL}}( p(\zb_{\leq k}|\zb_{>k}) q(\zb_{>k}|\xb) || p(\zb|\xb))
\end{equation}\label{eq:llr-as-kls-iwae-reduced}
which, in practice, is well-approximated for a finite $S$. We expect this importance weighted likelihood ratio to monotonically improve upon the one in \cref{eq:llr-as-kls} as $S$ increases and the KL-divergence in the regular ELBO that contains terms for which $\zb_i$ is high-dimensional goes to zero.

Since the scores in \cref{eq:llr-as-kls,eq:llr-as-kls-iwae-reduced} are estimated by sampling their estimators are stochastic objects with nonzero variance.
We note that $\text{Var}(\widehat{LLR}^{>k}) = \text{Var}(\hat{\mathcal{L}}) + \text{Var}(\hat{\mathcal{L}}^{>k}) - 2\, \text{Cov}(\hat{\mathcal{L}}, \hat{\mathcal{L}}^{>k})$.
Since $\log p_\thetab(\xb)$ and part of the KL divergence are identical in the expressions of $\mathcal{L}$ and $\mathcal{L}^{>k}$ we expect $\text{Cov}(\hat{\mathcal{L}}, \hat{\mathcal{L}}^{>k})$ to be positive which reduces the total variance. 
Empirical results indeed show that $\text{Var}(\widehat{LLR}^{>k})$ is larger than $\text{Var}(\hat{\mathcal{L}})$ but smaller than $\text{Var}(\hat{\mathcal{L}}^{>k})$.
%Whether this decreases the variance to below the variance of any of the two bounds, we do not know, and we believe it to be difficult to verify mathematically.
Nevertheless, the variance of the estimators is guaranteed to go to zero as the number of samples is increased.

@@

% Our model
%\begin{wrapfigure}{R}{0.5\columnwidth}
% \begin{SCfigure}[50][t!]
\begin{figure}[t]
    %\begin{figure}
    \centering
    \tikz{
        % inference
        % nodes
        \node[obs] (x_inf) {$\xb$};%
        \node[latent,above=.75cm of x_inf](z1_inf){$\zb_1$}; %
        \node[latent,above=.75cm of z1_inf](z2_inf){$\zb_2$}; %
        % \node[latent,above=.75cm of z2_inf](z3_inf){$\zb_3$}; %
        \node[above=of z2_inf, yshift=-1.cm] (phi) {$q_\phib(\zb|\xb)$};

@@

        % edges
        % \edge[]{z3_gen}{z2_gen};
        \edge[]{z2_gen}{z1_gen};
        \edge[]{z1_gen}{x_gen};
        \edge[dashed, bend left]{z2_gen}{x_gen};
        % \edge[dashed, bend left]{z3_gen}{x_gen};
    }
    \caption[Inference and generative models for a bottom-up hierarchical VAS]{The inference and generative models, $q_\phib$ and $p_\thetab$, for an $L=2$ layered bottom-up hierarchical VAE as the one used in our experiments.
    Dashed lines indicate deterministic skip connections which are employed in both networks. Skip connections are found to be useful for optimizing latent variable models \parencite{dieng_avoiding_2019, maaloe_biva_2019}.}
    \label{fig:hvae-graphical-model}
    % \end{figure}
%\end{wrapfigure}
% \end{SCfigure}
\end{figure}

@@

\begin{table}[t!]
    \centering
    \resizebox{0.8\columnwidth}{!}{%
    \begin{tabular}{lrrrrr}
        \toprule
         Method & Dataset & \multicolumn{4}{c}{Avg. bits/dim}\\
        %   & & $\log p(x)$ & $\mathcal{L}^{>k_1}$ & $\mathcal{L}^{>k_2}$ & $\mathcal{L}^{>k_3}$\\
          & & $\log p(x)$ & $\mathcal{L}^{>1}$ & $\mathcal{L}^{>2}$ & $\mathcal{L}^{>3}$\\
         \midrule
         \multicolumn{6}{c}{\textbf{Trained on FashionMNIST}} \\
         \midrule
         \multirow{2}{*}{Glow}
            & FashionMNIST & 2.96 & - & - & \\
            & MNIST & 1.83 & - & - & \\
         \multirow{2}{*}{HVAE (Ours)}
            & FashionMNIST & 0.420 & 0.476 & 0.579 & - \\
            & MNIST & 0.317 & 0.601 & 0.881 & - \\
         \midrule
         \multicolumn{6}{c}{\textbf{Trained on CIFAR10}} \\
         \midrule
         \multirow{2}{*}{Glow}
          & CIFAR10 & 3.46 & - & - & \\
          & SVHN & 2.39 & - & - & \\
         \multirow{2}{*}{HVAE (Ours)}
            & CIFAR10 & 3.74 & 17.8 & 54.3 & 75.7 \\  % log p(x) = 8010.03
            & SVHN & 2.62 & 10.2 & 64.0 & 93.9 \\
         \multirow{2}{*}{BIVA (Ours)}
          & CIFAR10 & 3.46 & 8.74 & 19.7 & 37.3 \\
          & SVHN & 2.35 & 6.62 & 25.1 & 59.0 \\
         \bottomrule
    \end{tabular}
    }
    \caption[Average bits per dimension of different datasets for generative models trained on FashionMNIST and CIFAR10]{%
    Average bits per dimension of different datasets for models trained on FashionMNIST and CIFAR10.
    For the hierarchical models we include the $\mathcal{L}^{>k}$ bounds.
    The likelihoods of training and test splits of the in-distribution data are all cases close.
    Since we train on dynamically binarized FashionMNIST, our bits/dim are smaller than for Glow.
    As $k$ is increased for the $L^{>k}$ bound, the bound gets looser but the model eventually assigns higher likelihood to the in distribution data than to the OOD data.
    Glow refers to \textcite{kingma_glow_2018, nalisnick_deep_2019}.
    BIVA refers to our implementation of \textcite{maaloe_biva_2019}.}
    \label{tab:bits-per-dim-ood}
    \vspace{0mm}
\end{table}

\footnotetext{%

@@

    Performance is worse when the overlap is high, cf. \textcite[Table 1]{serra_input_2020}, as seen with complex images.}

### 2023-04-12 chapters/paper_review/01_introduction.tex
Update reference keys for review paper

Representation learning approaches are generally considered examples of \textit{unsupervised
learning}, which refers to the family of machine learning methods that discover
naturally occurring patterns in training samples for which there are no pre-assigned
labels or scores~\cite{jordan_machine_2015}. 
The term ``unsupervised'' is used to distinguish this family of methods from
``supervised'' approaches, which assign a label to each training sample, and
``semi-supervised'' approaches, which utilize a small number of training samples
with labels to guide learning using a larger volume of unlabeled samples.
Examples of unsupervised learning techniques include k-means clustering~\cite{gray_vector_1984}, mixture models~\cite{jordan_hierarchical_1994}, autoencoders~\cite{hinton_autoencoders_1994},
and non-negative matrix factorization~\cite{lee_learning_1999}. 
\textit{Self-supervised learning} (SSL) is a fast-growing subcategory of
unsupervised learning approaches, which are techniques that utilize
information extracted from the input data itself as the label to learn
representations useful for 
downstream tasks. {\color{black} For example, unsupervised k-means clustering doesn't adhere to this definition of self-supervision since it iteratively minimizes the within-cluster variance during learning.}
In this review, we
focus on self-supervised learning approaches.

>>>>

Representation learning approaches are generally considered examples of \textit{unsupervised
learning}, which refers to the family of machine learning methods that discover
naturally occurring patterns in training samples for which there are no pre-assigned
labels or scores~\cite{jordan_machine_2015}. 
The term ``unsupervised'' is used to distinguish this family of methods from
``supervised'' approaches, which assign a label to each training sample, and
``semi-supervised'' approaches, which utilize a small number of training samples
with labels to guide learning using a larger volume of unlabeled samples.
Examples of unsupervised learning techniques include $k$-means clustering~\cite{gray_vector_1984}, mixture models~\cite{jordan_hierarchical_1994}, autoencoders~\cite{hinton_autoencoders_1994},
and non-negative matrix factorization~\cite{lee_learning_1999}. 
\textit{Self-supervised learning} (SSL) is a fast-growing subcategory of
unsupervised learning approaches, which are techniques that utilize
information extracted from the input data itself as the label to learn
representations useful for 
downstream tasks. {\color{black} For example, unsupervised $k$-means clustering doesn't adhere to this definition of self-supervision since it iteratively minimizes the within-cluster variance during learning.}
In this review, we
focus on self-supervised learning approaches.

### 2023-04-12 chapters/paper_review/02_three_waves.tex
Update reference keys for review paper

Early work used simple clustering methods. For example, in work such
as \cite{rabiner_considerations_1979,wilpon_modified_1985}, word patterns were clustered
semi-automatically using techniques such as k-means, after which isolated words
were recognized by finding the training cluster closest to 
  the test data.

@@

In the SSL context, a pretext task related to autoencoding is to generate
an object from its partial information. Such tasks are widely used in NLP, for
example, using 
the previous tokens 
in a sentence to predict the next token such as in
ELMo~\cite{peters_deep_2018}, the GPT
series~\cite{radford_language_2019}, and 
Megatron~\cite{shoeybi_megatronlm_2020}, or predicting the masked tokens in a
sentence such as with the bidirectional encoder representations from Transformers
(BERT) series~\cite{devlin_bert:_2018,liu_roberta_2019}. 
Another common pretext task in the third wave is contrastive
learning~\cite{oord_representation_2018}, in which a model learns to identify a
target instance from a set of negative samples. 
This approach has become especially popular in the
CV context~\cite{chen_simple_2020,he_momentum_2020,chen_improved_2020,caron_unsupervised_2020}.
In this survey, we will mainly focus on techniques for pretext task
optimization for speech processing, and discuss these techniques in detail
in \cref{sec:approach}.

>>>>

Early work used simple clustering methods. For example, in work such
as \cite{rabiner_considerations_1979,wilpon_modified_1985}, word patterns were clustered
semi-automatically using techniques such as $k$-means, after which isolated words
were recognized by finding the training cluster closest to 
  the test data.

@@

In the SSL context, a pretext task related to autoencoding is to generate
an object from its partial information. Such tasks are widely used in NLP, for
example, using 
the previous tokens 
in a sentence to predict the next token such as in
ELMo~\cite{peters_deep_2018}, the GPT
series~\cite{radford_language_2019}, and 
Megatron~\cite{shoeybi_megatronlm_2020}, or predicting the masked tokens in a
sentence such as with the bidirectional encoder representations from Transformers
(BERT) series~\cite{devlin_bert_2018,liu_roberta_2019}. 
Another common pretext task in the third wave is contrastive
learning~\cite{oord_representation_2018}, in which a model learns to identify a
target instance from a set of negative samples. 
This approach has become especially popular in the
CV context~\cite{chen_simple_2020,he_momentum_2020,chen_improved_2020,caron_unsupervised_2020}.
In this survey, we will mainly focus on techniques for pretext task
optimization for speech processing, and discuss these techniques in detail
in \cref{sec:approach}.
